## Model

```{r}
mlp_mod <- 
  mlp(
    mode = "classification",
    hidden_units = tune(),
    dropout = tune(),
    epochs = 50,
  ) |> 
  set_engine("keras", verbose = 0)
mlp_mod
```

## Workflow in R and pipeline in Python

Here we bundle the preprocessing step (Python) or recipe (R) and model.

```{r}
mlp_wf <- workflow() |>
  add_recipe(mlp_recipe) |>
  add_model(mlp_mod)
mlp_wf
```

## Tuning grid

Here we tune the `cost` and radial scale `rbf_sigma`.

```{r}
param_grid <- grid_regular(
  hidden_units(range = c(1, 20)),
  dropout(range = c(0, 0.6)),
  levels = 5
  )
param_grid
```


## Cross-validation (CV)

::: {.panel-tabset}

#### R

Set cross-validation partitions.
```{r}
set.seed(203)

folds <- vfold_cv(mimic_other, v = 5)
folds
```

Fit cross-validation.
```{r}


mlp_fit <- mlp_wf |>
  tune_grid(
    resamples = folds,
    grid = param_grid,
    metrics = metric_set(roc_auc, accuracy)
    )
mlp_fit
```

Visualize CV results:
```{r}
mlp_fit |>
  collect_metrics() |>
  print(width = Inf) |>
  filter(.metric == "roc_auc") |>
  ggplot(mapping = aes(x = dropout, y = mean, color = factor(hidden_units))) +
  geom_point() +
  labs(x = "Dropout Rate", y = "CV AUC") +
  scale_x_log10()
```

Show the top 5 models.
```{r}
mlp_fit |>
  show_best("roc_auc")
```
Let's select the best model.
```{r}
best_mlp <- mlp_fit |>
  select_best("roc_auc")
best_mlp
```

## Finalize our model

Now we are done tuning. Finally, let’s fit this final model to the whole training data and use our test data to estimate the model performance we expect to see with new data.

```{r}
# Final workflow
final_wf <- mlp_wf |>
  finalize_workflow(best_mlp)
final_wf
```

```{r}
# Fit the whole training set, then predict the test cases
final_fit <- 
  final_wf |>
  last_fit(data_split)
final_fit
```

```{r}
# Test metrics
final_fit |> 
  collect_metrics()
```
